Manuscript

### 2025-03-19 Manuscript_MD_ScoreReLiability.qmd
start on second re-analysis

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the Shnabel_Willingness_Reconcile_Rev, the
reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available. Again,
only for the Srull_Behaviour_Hostility phenomenon, we find a substantial
reduction in heterogeneity due to the attenuation correction procedure.
However, for the majority of phenomena it appears that differences in
score reliability do not inflate heterogeneity in uncorrected effect
sizes, but instead masks heterogeneity, that can be unveiled by
accounting for score reliability.

@@

@tbl-tauES showed that for all phenomena, where we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Subsequently, in
Equation 16 we demonstrate that if the $R_2$ metric is larger than
$R_1^3$, ES heterogeneity is bound to be larger after applying said
procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
true score variances, including their heterogeneity and metrics $R_1$
and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across nine phenomena, $R_2$ is larger
than .97. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution. Next to the two exceptional phenomena,
where zero heterogeneity in score variance was found
(Alter_Analytic_Processing and Shnabel_Willingness_Reconcile_Rev), only
for the Srull_Ronald_Hostility phenomenon some heterogeneity could be
attributed to differences in random error variance, albeit less than
14%.

This implies that for all measurements concerning these phenomena, the
heterogeneity in observed score variances could not be substantially
explained by differences in measurement quality. For no phenomenon, if
differences in score variance could be found, was $R_2$ smaller than
$R_1^3$. As demonstrated in Equation 20, this implies that the ES
heterogeneity grows even larger, as we correct for differences in score
reliability.

@@

In @tbl-tauES we also found some phenomena where the heterogeneity did
in fact shrink after applying the attenuation correction procedure. This
is not reflected by what we observe in @tbl-R1R2, as $R_2 > R_1$ for
almost all phenomena, which would imply that across all phenomena we
expected ES heterogeneity to grow as a consequence of the attenuation
correction procedure. However, it is crucial to realise that we did not
find statistically significant ES heterogeneity for any of the phenomena
where ES heterogeneity was reduced. This implies, that we actually could
not statistically distinguish the ES heterogeneity from zero. Therefore,
what seems to be a reduction in ES heterogeneity was most likely caused
by estimation issues, as we could not accurately estimate the extent of
ES heterogeneity, either before or after the attenuation correction.

Lastly, for two out of the 12 phenomena, we find a $\tau_{\sigma^2_X}$
of 0. For these phenomena, it is therefore also impossible to compute
the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
did not identify statistically significant ES heterogeneity in the first
place, we did find some for the Shnabel_Willingness_Reconcile_Rev
phenomenon. At the same time, we did not identify statistically
significant heterogeneity in score reliability. With zero heterogeneity
in score variances, Equation 13 implies that the ES heterogeneity is
bound to grow larger due to the attenuation procedure. Also, this
phenomenon was assessed with a particularly low number of replications
(8). Most likely, adequate power to detect heterogeneity in score
reliability and observed score variance requires a larger number of
replications than adequate power to detect heterogeneity. In that case,
the results from Table 5, concerning the
Shnabel_Willingness_Reconcile_Rev phenomenon might alternatively be
explained by power-issues. Similarly, while Table 3 does not imply
specific expectations concerning the presence of score variance
heterogeneity in the Alter_Analytic_Processing phenomenon, the low
number of replications (20) makes it hard to distinguish whether the
estimate of $\tau_{\sigma^2_X} = 0$ is actually sensible or can also be
attributed to low power.

>>>>

Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
heterogeneity grows as a result from the attenuation correction
procedure. Even though these differences tend to be small, it is clear
that in all those cases, not only did accounting for score reliability
not explain any ES heterogeneity, we find even more than initially
expected. This is incompatible with claims made in W&D and H&S, the
attenuation correction increased ES heterogeneity. Concerning the
remaining two phenomena, for the Shnabel_Willingness_Reconcile_Rev, the
reduction in heterogeneity is miniscule and the estimate of
heterogeneity is not statistically significant in the first place. Most
likely, the difference in ES heterogeneity estimates can be explained by
issues of estimation, as only eight replications are available.
Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
substantial reduction in heterogeneity due to the attenuation correction
procedure. The estimates of ES heterogeneity are not statistically
significant, however, which hampers our ability to meaningfully
interpret and compare these estimates. For the majority of phenomena it
appears that differences in score reliability do not inflate
heterogeneity in uncorrected effect sizes, but instead mask
heterogeneity, that can be unveiled by accounting for score reliability.

@@

@tbl-tauES showed that for all phenomena, where we identified
statistically significant ES heterogeneity, the heterogeneity was even
larger after the attenuation correction procedure. Subsequently, in
Appendix A we demonstrate that if the $R_2$ metric is larger than
$R_1^3$, ES heterogeneity is bound to be larger after applying said
procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
true score variances, including their heterogeneity and metrics $R_1$
and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
Additionally, we observe that across nine phenomena, $R_2$ is larger
than .96. This implies that of the heterogeneity observed in score
variances, 97% or more could not be explained by differences in random
error introduced by the measuring instrument, but by genuine differences
in the underlying distribution.

For one of the phenomena, however, we observe a large disparity compared
to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
phenomenon, $R_2$ is very low, below .25. This makes it even lower than
its $R_1^3 = .574$. This would imply that for this phenomenon, the
conditions for a reduction in ES heterogeneity would be met. However, in
@tbl-tauES, we find no such reduction. What becomes apparent, however,
is that for this phenomenon we are working with the low number of 9
administration sites. Estimation precision for heterogeneity in variance
components, even more so than for ES, largely depends on the sample
size. Beinhauer et al. (ref) come to the conclusion that at 10 items,
which the WIllingness to Reconcile scale consists of, a sample size
around 33 is required to achieve sufficient power for a statistical test
of score variance heterogeneity. Most likely, the low size led to
implausible estimates in both true and total score variance
heterogeneity and subsequently to an implausible estimate of $R_2$.
Meta-analytic estimates, especially those concerning score variance
components such as $R_2$, therefore should be interpreted with caution,
if the sample size was particularly low.

Similarly, for the two exceptional phenomena where the attenuation
correction procedure actually led to an increase in ES heterogeneity
(Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
significance test for the ES heterogeneity was not significant, both
before and after the correction. The results regarding $R_2$ indicate
that the majority of score variance heterogeneity was due to differences
in true score variance $\sigma^2_T$. Both phenomena were examined using
a rather low number of administration sites (22 and 9 respectively),
implying that issues of estimation precision might explain the
inconclusive results. Lastly, for one out of the 12 phenomena, the
Albarracin_Priming_SAT phenomenon, we find a $\tau_{\sigma^2_X}$ of 0.
Therefore, the $R_2$ metric could not be computed, as the denominator
would have been zero. However, we also did not identify statistically
significant ES heterogeneity in the first place. This phenomenon was
assessed with a particularly low number of replications (8). As
discussed, adequate power to detect heterogeneity in score reliability
and observed score variance requires a larger number of replications
than adequate power to detect heterogeneity.

Generally, @tbl-tauES implies that for almmost all measurements
concerning these phenomena, the heterogeneity in observed score
variances could not be substantially explained by differences in
measurement quality. Only for one phenomenon, where differences in score
variance could be found, $R_2$ was smaller than $R_1^3$. As discussed,
for this phenomenon, the number of administration sites was too low to
accurately estimate heterogeneity in score variance components. For all
remaining phenomena, $R_2$ was larger than $R_1$, implying that an
attenuation correction procedure could not reduce ES heterogeneity. The
results in @tbl-tauES agree with this conclusion, as no phenomena could
be identified where statistically significant ES heterogeneity was
accompanied by a reduction in differences due to the correction.

@@
